Algebraic File Synchronization: Adequacy and Completeness

### 2018-02-05 p2/axioms.tex
Converted axioms

% TODO \footnote{The program is accessible on-line at \\
% https://github.com/csirmaz/AlgebraicSyncPaper/blob/master/p2/prove.py.}

@@

\begin{mydef}[Construction and destruction pairs]
A pair of commands on nodes $\nn$ and $n$ is a construction pair if $\nn=\parentf{n}$ 
and the commands are one
of the following:
   \begin{align*}
            \cbda{\nn}&\cc  \cbfa{n} \\
            \cbda{\nn}&\cc  \cbda{n} \\
            \cfda{\nn}&\cc  \caaa{\empt}{\valff}{n} \\
            \cfda{\nn}&\cc  \cbda{n}
   \end{align*}
\end{mydef}

\begin{mydef}[Destruction pair]
A pair of commands on nodes $n$ and $\nn$ is a destruction pair if $\parentf{n}=\nn$ and the commands are one
of the following:
   \begin{align*}
            \cfba{n}&\cc  \cdba{\nn} \\
            \cfba{n}&\cc  \caaa{\vald}{\valff}{\nn} \\
            \cdba{n}&\cc  \cdba{\nn} \\
            \cdba{n}&\cc  \cdfa{\nn}
   \end{align*}
\end{mydef}

\begin{myax}\mlabel{ax_separate_commute}
Commands on incomparable nodes commute:
$\cxynv\cc\czwmv \equiv \czwmv\cc\cxynv$ where $n\unrel m$.
\end{myax}

\begin{myax}\mlabel{ax_separate_nobreaks}
Commands on incomparable nodes do not break all filesystems:
$\wrksx{\cxynv\cc\czwmv}$ where $n\unrel m$.
\end{myax}

\begin{myax}\mlabel{ax_same_breaks}
Commands on the same node break every filesystem if their types are incompatible:
$\cxynv\cc\czwnv \equiv \cbrk$ where $[y]\ne [z]$.
\end{myax}

\begin{myax}\mlabel{ax_same_emptyseq}
Commands on the same node simplify:
$\cxynv\cc\czwnv \eqext \emptyseq$ where $[y]=[z]$, and $\cxw=\cbb$ 
or $\cxw=\cdd$.
\end{myax}

\begin{myax}\mlabel{ax_same_singlec}
Commands on the same node simplify:
$\cxynv\cc \czwnv \equiv \cxwnv$ where $[y]=[z]$ and $\cxw\neq\cbb$ and $\cxw\neq\cdd$.
\end{myax}

\begin{myax}\mlabel{ax_distantrel_breaks}
Commands on distant relatives break all filesystems:
$\cxynnv\cc\czwnv \equiv \cbrk$
and $\czwnv\cc\cxynnv \equiv\cbrk$
where $\nn\descendant n$ and $\nn\neq\parent(n)$, and $\cxy\neq\cdd$ and $\czw\neq\cbb$.
(More intuitively, if $\czw\neq\cbb$ then either before or after $\czwnv$, the filesystem at
$n$ is not empty, and therefore there must be a directory at all ancestors of $n$, including at a child of $\nn$.
Consequently $\cxynnv$ will break the filesystem as either before or after it the filesystem at $\nn$
does not have a directory.)
\end{myax}

\begin{myax}\mlabel{ax_directchild_breaks}
All other commands on a child break every filesystem:
$\cxynnv\cc\czwnv \equiv \cbrk$ where $\nn=\parent(n)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$
and the pair is not a construction pair.
As we shall see, this can be extended to \cref{simple_distant_pairs}.
TODO TO BE DELETED -- WE ONLY WANT TO USE \cref{simple_distant_pairs}
\end{myax}

\begin{myax}\mlabel{ax_directparent_breaks}
All other commands on a parent break every filesystem:
$\cxynv\cc\czwnnv \equiv \cbrk$ where $\parent(n)=\nn$ and $\cxy\neq\cbb$ and $\czw\neq\cdd$
and the pair is not a destruction pair.
This can also be extended to \cref{simple_distant_pairs}.
TODO TO BE DELETED -- WE ONLY WANT TO USE \cref{simple_distant_pairs}
\end{myax}

\begin{mydef}[Assertion command]
A command is an assertion command if
it leaves every filesystem in the same state
where it is defined.
In other words, assertion commands have the same input and output type,
and there must also be only one possible value for their output.
Accordingly, $\cbba{n}$ and $\cdda{n}$ are the only two types of assertion commands.
\end{mydef}

\begin{myax}\mlabel{ax_child_assert}
An assertion command can be added on a descendant node:
\[ \cbba{n}\cc\cxynnv \equiv \cxynnv \equiv \cxynnv\cc\cbba{n} \] 
where $\nn\descendant n$ and $\cxy\neq\cdd$.
(Note that this and next rule is true because the assertion command is
adjacent to the original command.)
\end{myax}

\begin{myax}\mlabel{ax_parent_assert}
An assertion command can be added on an ancestor node:
\[ \cdda{\nn}\cc\cxynv \equiv \cxynv \equiv \cxynv\cc\cdda{\nn} \]
where $\nn\descendant n$ and $\cxy\neq\cbb$.
\end{myax}

\begin{myax}\mlabel{ax_assert}
Assertion commands can be removed:
$\cdda{n}\eqext\emptyseq$ and $\cbba{n}\eqext\emptyseq$.
% $\cxynv \eqext \emptyseq$ where $\cxy=\cbb$ or $\cxy=\cdd$.
\end{myax}

>>>>

@@

\begin{mydef}[Construction and destruction pairs]
A pair of commands on nodes $\nn$ and $n$ where $\nn=\parentf{n}$
is a \emph{construction pair} if the input and output types match
one of the following patterns:
   \begin{align*}
            \cbda{\nn}&\cc  \cbfa{n} \\
            \cbda{\nn}&\cc  \cbda{n} \\
            \cfda{\nn}&\cc  \caaa{\empt}{\valff}{n} \\
            \cfda{\nn}&\cc  \cbda{n}
   \end{align*}
The pair is a \emph{destruction pair} if the types match one of the following:
   \begin{align*}
            \cfba{n}&\cc  \cdba{\nn} \\
            \cfba{n}&\cc  \caaa{\vald}{\valff}{\nn} \\
            \cdba{n}&\cc  \cdba{\nn} \\
            \cdba{n}&\cc  \cdfa{\nn}
   \end{align*}
\end{mydef}

We can see that construction pairs consist of construction comands,
while destruction pairs consist of destruction commands.

\Cref{rules_lemma} summarizes the basic properties of command pairs.
The claims listed are named \emph{\namecref{ax_separate_commute}}
as they can also be interpreted as inference rules in a pure
algebraic treatment of filesystem synchronization.
TODO elaborates this approach.

\begin{mylem}\mlabel{rules_lemma}
\newcounter{rulecounter}
\begin{list}{\bf Rule~\arabic{rulecounter}.}{\usecounter{rulecounter}}

\item[] % line break after the lemma label

\item\mlabel{ax_separate_commute}
Commands on incomparable nodes commute:
$\cxynv\cc\czwmv \equiv \czwmv\cc\cxynv$ where $n\unrel m$.

\item\mlabel{ax_separate_nobreaks}
Commands on incomparable nodes also do not break all filesystems:
$\wrksx{\cxynv\cc\czwmv}$ where $n\unrel m$.

\item\mlabel{ax_same_breaks}
Commands on the same node break every filesystem if their types are incompatible:
$\cxynv\cc\czwnv \equiv \cbrk$ where $[y]\ne [z]$.

\item\mlabel{ax_same_emptyseq}
Otherwise commands on the same node are extended by an empty sequence if their outer types represent
an assertion command:
$\cxynv\cc\czwnv \eqext \emptyseq$ where $[y]=[z]$, and $\cxw=\cbb$ 
or $\cxw=\cdd$.

\item\mlabel{ax_same_singlec}
All other command pairs on the same node are equivalent to a single command:
$\cxynv\cc \czwnv \equiv \cxwnv$ where $[y]=[z]$ and $\cxw\neq\cbb$ and $\cxw\neq\cdd$.

% TODO Renumber rules in script

\item\mlabel{ax_directchild_breaks}
Commands on a parent and a child node break every filesystem
if the pair is not a construction pair
and the commands do not simply assert a directory at the parent
or an empty node at the child:
$\cxynnv\cc\czwnv \equiv \cbrk$ where $\nn=\parent(n)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$.

\item\mlabel{ax_directparent_breaks}
Commands on a child and parent node break every filesystem
if the pair is not a destruction pair
and the commands do not simply assert an empty node at the child
or a directory at the parent:
$\cxynv\cc\czwnnv \equiv \cbrk$ where $\parent(n)=\nn$ and $\cxy\neq\cbb$ and $\czw\neq\cdd$.

\item\mlabel{ax_distantrel_breaks}
Commands on distant relatives break all filesystems
if the commands do not simply assert a directory at the ancestor
or an empty node at the descendant:
$\cxynnv\cc\czwnv \equiv \cbrk$
and $\czwnv\cc\cxynnv \equiv\cbrk$
where $\nn\descendant n$ and $\nn\neq\parent(n)$, and $\cxy\neq\cdd$ and $\czw\neq\cbb$.

\item\mlabel{ax_child_assert}
An assertion command can be freely added on a descendant node
next to a command that does not simply assert a directory:
\[ \cbba{n}\cc\cxynnv \equiv \cxynnv \equiv \cxynnv\cc\cbba{n} \] 
where $\nn\descendant n$ and $\cxy\neq\cdd$.

\item\mlabel{ax_parent_assert}
An assertion command can be freely added on an ancestor node
next to a command that does not simply assert an empty node:
\[ \cdda{\nn}\cc\cxynv \equiv \cxynv \equiv \cxynv\cc\cdda{\nn} \]
where $\nn\descendant n$ and $\cxy\neq\cbb$.

\item\mlabel{ax_assert}
All assertion commands can be removed as the empty sequence extends them:
$\cdda{n}\eqext\emptyseq$ and $\cbba{n}\eqext\emptyseq$.

\end{list}
\end{mylem}

These claims are easy but tedious to check.
To ensure that all cases were considered when verifying these claims,
a computer program was prepared that models the relevant parts of
a filesystem and checks the results of applying all different command pairs
against the above claims.\footnote{The program is accessible at \\ 
https://github.com/csirmaz/AlgebraicSyncPaper/blob/master/p2/prove.py.}

For \cref{ax_distantrel_breaks}, we also provide a simple proof.
Suppose $\cxynnv$ and $\czwnv$ satisfy the conditions in the \namecref{ax_distantrel_breaks},
and there if an $\FS$ for which $(\cxynnv\cc\czwnv)\aFS\neq\fsbroken$.
If $\czw\neq\cbb$, then either $\FS(n)\neq\empty$ or $\czwnv\aFS(n)\neq\empty$,
and therefore $\FS(\parentf{n})$ must be a directory for the filesystem to satisfy
the tree property at all times.
However, from $\cxy\neq\cdd$ we also know that either $\FS(\nn)$ 
(before applying $\cxynv$) or $\cxynv\aFS(\nn)$ (after applying $\cxynv$)
is not a directory. Depending on which is true, we have a point in the sequence
of applying the commands where the filesystem cannot satify the tree property
and therefore becomes broken as $\parentf{n}$, a descendant of $\nn$, contains
a directory and so must all its ancestors.

For \cref{ax_child_assert,ax_parent_assert}, we note that these claims
easily follow from the fact that the commands are adjacent to each other.

Finally, we have included \cref{ax_directchild_breaks,ax_directparent_breaks}
for completeness, but we will use the more generic \cref{simple_distant_pairs},
which extends these rules to non-adjacent commands, although only in the case
of command sequences without any assertion commands.

### 2018-02-05 p2/def-filesystem.tex
Command categories
Updated script
Converted axioms

Every filesystem has a so-called \emph{tree property}, which means that
if the filesystem is not empty at a node, and the node has a parent,
then there must be a directory at the parent node.
Using the notation introduced above, we can formally express this as follows.
\begin{mydef}[Tree property]
A filesystem $\FS:\setn \rightarrow \setv$ has the tree property iff
\[ \forall n\in\setn
\FS(n) \neq \empt \Rightarrow \FS(\parent(n)) \typeeq \vald \]
wherever $\parent(n)$ is defined, that is, $n$ has a parent.
\end{mydef}

@@

\smallskip
In the rest of this paper,
we fix the skeleton $\langle\setn,\parent\rangle$,
and the set of values $\setv$.
$\FS$ (with or without indices) denotes a filesystem,
and $n$, $m$ and $o$ are nodes in $\setn$,
and we write $\nn$ for the parent of $n$, assuming it exists.

>>>>

Every filesystem has a so-called \emph{tree property}, which means that
if the filesystem is not empty at a node, and the node has a parent,
then there must be a directory at the parent node.
Using the notation introduced above, we can formally express this as follows.
\begin{mydef}[Tree property]
A filesystem $\FS:\setn \rightarrow \setv$ has the tree property iff
\[ \forall n\in\setn
\FS(n) \neq \empt \Rightarrow \FS(\parentf{n}) \typeeq \vald \]
wherever $\parent(n)$ is defined, that is, $n$ has a parent.
\end{mydef}

@@

\myskip
In the rest of this paper,
we fix the skeleton $\langle\setn,\parent\rangle$,
and the set of values $\setv$.
$\FS$ (with or without indices) denotes a filesystem,
and $n$, $m$ and $o$ are nodes in $\setn$,
and we write $\nn$ for the parent of $n$, assuming it exists.
